
#Loading all libraries


```{r}

install.packages("gganimate")
library(gganimate)
install.packages("stringr")
install.packages("tidyverse")
install.packages("fpp2")
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(lubridate)
library(fpp2)
library(reshape2)
library(hms)
library(stringr)
library(tidyr)
install.packages('devtools')
install.packages("ggalt")

devtools::install_github('bbc/bbplot')
install.packages("pacman")
library(bbplot)
library(ggalt)


install.packages("treemap")
library(treemap)

install.packages("readxl")
library(readxl)

```

Importing all the datasets
```{r}

#This file normally takes 3minutes to import

p_2019 <- read.csv("E:/DAE/FIRST SEM/Computation and Visualization/datasets/nyc-parking-tickets/Parking_Violations_Issued_-_Fiscal_Year_2019.csv", na.strings="", stringsAsFactors=FALSE)

colnames(p_2019)[1]<- "Summons.Number"

vcode <- read_excel("E:/DAE/FIRST SEM/Computation and Visualization/datasets/nyc-parking-tickets/ParkingViolationCodes_Nov_2018.xlsx")

colnames(vcode)=c('Violation.Code', 'desc', 'fine_man','fine_other')



```
Doing changes to the  Fiscal Year 2019 data

```{r}


#Calculating hour of the day and the day of the week
#*******************************************************


p_2019$Issue.Date <- as.Date(p_2019$Issue.Date, format = "%m/%d/%Y")

p_2019$Day <- wday(p_2019$Issue.Date, label = TRUE) 

p_2019 <-mutate(p_2019, Hour = ifelse(str_detect(Violation.Time,'A'), substr(Violation.Time,1,2), as.numeric(as.numeric(substr(Violation.Time,1,2))+12 )))

p_2019 <- mutate(p_2019, Hour = ifelse( Hour=='12','00',ifelse(Hour=='24','12',Hour)))

#Merging the dataframe with vcode

p_2019<-merge(p_2019, vcode, by = 'Violation.Code')

#Getting the quarter 


p_2019$Quarter <- quarter(p_2019$Issue.Date)

```
b) Adding the borough column
```{r}
p_2019 <- mutate(p_2019, Borough =ifelse(Violation.County == "K"|Violation.County == "KINGS"|Violation.County == "KING"|Violation.County == "BK", "Brooklyn", ifelse(Violation.County == "Q" |Violation.County == "QU"| Violation.County == "QUEEN"|Violation.County == "QN"|Violation.County == "QNS", "Queens",ifelse(Violation.County == "R"| Violation.County == "Rich"|Violation.County == "ST", "Staten Island",ifelse(Violation.County == "NY"|Violation.County == "NEWY"|Violation.County == "MAN"|Violation.County == "MH"|Violation.County == "NEW Y"|Violation.County == "MN", "Manhattan", ifelse(Violation.County == "BX"|Violation.County == "PBX"|Violation.County == "BRONX", "Bronx", "Invalid"))))  ))

#Removing rows with invalid values in Borough Column
p_2019 <- filter(p_2019,p_2019$Borough!= "Invalid")


```





Plot to show the peak hours during the week

```{r,warning=FALSE}
day_hour<- p_2019%>%
  group_by(Day,Hour)%>%
  summarise(total=n())

day_hour <- filter(day_hour, Hour %in% c('06','07','08','10','11','12','13','14','15','16','17','18','19','20','21','22','23') )



week_hour_analysis <- ggplot(day_hour, aes(Hour,factor(Day)))+ 
  geom_tile(aes(fill= total))+
  ylab(" Day of the week")+
  xlab("Hours")+ labs(fill= "Number of parking violations") + ggtitle("Peak hours during the week")+scale_fill_gradient(low="blue", high="red")+
  theme(plot.title = element_text(family="Times New Roman", face="bold", size=18), axis.title.x = element_text(family="Arial",size=16), axis.title.y = element_text(family="Arial",size=16), legend.title = element_text(family="Arial",size=16))

week_hour_analysis
```


Finding which violations decreased the most in Quarter 2

```{r,warning=FALSE}


quarter_analysis <- p_2019%>%
  filter(Quarter=='1'| Quarter =='2')%>%
  group_by(desc, Quarter)%>%
  summarise(total=n())



dumbbell_dataframe <- quarter_analysis %>%
  select(desc, Quarter, total) %>%
  spread(Quarter, total) 
  
  #Changing NA to 0
  dumbbell_dataframe[is.na(dumbbell_dataframe)]<-0  


#Creating the final df for plot  
dumbbell_final <- dumbbell_dataframe%>%
  mutate(gap = `1` - `2`) %>%
  arrange(desc(gap)) %>%
  head(10)

#Plotting the dumbell plot



ggplot(dumbbell_final, aes(x = `1`, xend = `2`, y = reorder(desc, gap), group = desc)) + 
  geom_dumbbell(colour = "#dddddd",
                size = 3,
                colour_x = "#FAAB18",
                colour_xend = "#1380A1",show.legend = TRUE) +
  bbc_style() + 
  labs(title="What happened in Quarter 2?",
       subtitle=" Decrease in violations, Quarter1-Quarter2")


```

Plotting the Treemap

```{r,warning=FALSE}
df = count(p_2019,Borough,Violation.Precinct)
df= filter(df,Violation.Precinct!="0")
treemap(df,index = c("Borough","Violation.Precinct"),vSize = "n",palette="Set3",border.col=c("black","white"),border.lwds=c(7,2),title = "Violation Treemap For Boroughs and Precincts",fontsize.labels = c(15,10),align.labels = list(c("centre","centre"),c("left","top")))
```

Creating the dataframe for tableau file to show distribution of revenue across precincts
```{r}

tableau_df <-  count(p_2019, Violation.Precinct, Violation.Code)
tableau_df= filter(tableau_df,Violation.Precinct!="0")  
#Adding fine to the violation code
tableau_df <- left_join(tableau_df, select(vcode, -c(desc,fine_other)),by="Violation.Code")
tableau_df <- mutate(tableau_df, revenue = n*fine_man)%>%
  group_by(Violation.Precinct)%>%
  summarise(total = sum(revenue))
 
  
write.csv(tableau_df, 'Tableau_file.csv')

  
```

Importing the data of 2017 and 2018 

```{r}
p_2017 <- read.csv("E:/DAE/FIRST SEM/Computation and Visualization/datasets/nyc-parking-tickets/Parking_Violations_Issued_-_Fiscal_Year_2017.csv", na.strings="", stringsAsFactors=FALSE)

p_2018 <- read.csv("E:/DAE/FIRST SEM/Computation and Visualization/datasets/nyc-parking-tickets/Parking_Violations_Issued_-_Fiscal_Year_2018.csv", na.strings="", stringsAsFactors=FALSE)

colnames(p_2018)[1]<- "Summons.Number"



#Changing Issue Date to Date format
p_2017$Issue.Date <- as.Date(p_2017$Issue.Date, format = "%m/%d/%Y")

p_2018$Issue.Date <- as.Date(p_2018$Issue.Date, format = "%m/%d/%Y")


#Selecting relevant columns
p_2017 <- select(p_2017,  Summons.Number,  Registration.State,Issue.Date,  Violation.Code,  Vehicle.Body.Type, Vehicle.Make , Vehicle.Expiration.Date, Violation.Precinct, Violation.Time, Vehicle.Year,Violation.County)

p_2018 <- select(p_2018,Summons.Number,  Registration.State,Issue.Date,  Violation.Code,  Vehicle.Body.Type, Vehicle.Make , Vehicle.Expiration.Date, Violation.Precinct, Violation.Time, Vehicle.Year, Violation.County)

p_2019 <- select(p_2019,Summons.Number,  Registration.State,Issue.Date,  Violation.Code,  Vehicle.Body.Type, Vehicle.Make , Vehicle.Expiration.Date, Violation.Precinct, Violation.Time, Vehicle.Year, Violation.County)




```

Creating the dataframe by merging the datasets

```{r}
p_merged <- rbind(p_2017,p_2018,p_2019)
```


Plotting the quarter wise analysis of number of violations
```{r,warning= FALSE}
  
p_merged$Quarter <- quarter(p_merged$Issue.Date)

p_merged$Year <- year(p_merged$Issue.Date)



time_analysis <- p_merged%>%
  group_by(Year, Quarter)%>%
  summarise(total= n())  
  
time_analysis <- filter(time_analysis, Year %in% c(2017,2018,2019))  
time_analysis <- time_analysis[-c(11,12),]

time_analysis$Label <- c("2017 Q1","2017 Q2","2017 Q3","2017 Q4","2018 Q1","2018 Q2","2018 Q3","2018 Q4","2019 Q1","2019 Q2")

#Plotting
  
ggplot(time_analysis, aes(x = Label, y = total,group =1, color= factor(Year))) + 
  geom_line(stat = 'identity')+
  geom_point()+
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  bbc_style() +
  labs(title="Parking Violations in NYC 2017-2019")
  
```

```{r}

p_2019$Quarter <- quarter(p_2019$Issue.Date)

p_2019 <- mutate(p_2019, Borough =ifelse(Violation.County == "K"|Violation.County == "KINGS"|Violation.County == "KING"|Violation.County == "BK", "Brooklyn", ifelse(Violation.County == "Q" |Violation.County == "QU"| Violation.County == "QUEEN"|Violation.County == "QN"|Violation.County == "QNS", "Queens",ifelse(Violation.County == "R"| Violation.County == "Rich"|Violation.County == "ST", "Staten Island",ifelse(Violation.County == "NY"|Violation.County == "NEWY"|Violation.County == "MAN"|Violation.County == "MH"|Violation.County == "NEW Y"|Violation.County == "MN", "Manhattan", ifelse(Violation.County == "BX"|Violation.County == "PBX"|Violation.County == "BRONX", "Bronx", "Invalid"))))  ))

#Removing rows with invalid values in Borough Column
p_2019 <- filter(p_2019,p_2019$Borough!= "Invalid")


```

Finding out which violations are being taken casuallu


```{r}

vcode_fine <- p_2019%>%
  filter(Borough=="Manhattan")%>%
  group_by(Violation.Code,fine_man,desc)%>%
  summarise(total=n())%>%
  arrange(desc(total))%>%
  head(20)%>%
  mutate(ratio=total/fine_man)%>%
  arrange(desc(ratio))%>%
  head(10)
  
```

Plotting the map

```{r,warning=FALSE}

ggplot(vcode_fine, aes(x=total, y=fine_man, color=factor(desc))) + 
  geom_point()+xlab('Number of violations')+ ylab("Fine")+ggtitle("Which violations are taken casually?")+labs(color='Violation Description') +
  theme(plot.title = element_text(family="Times New Roman", face="bold", size=18), axis.title.x = element_text(family="Arial",size=16), axis.title.y = element_text(family="Arial",size=16), legend.title = element_text(family="Arial",size=16))
  



```



